### 2016-07-20 rebuttal_reviewers.txt

First, we thank the reviewers for their thoughtful commentary on our paper to keep the quality of
NIPS top-notch.

General comments:

We have been working continuously on the paper in the last month, and have been able to address a
number of our reviewer comments. For instance, we have been able to demonstrate worst-case scenarios
for references [3] and [4] with Omega(N) lower bounds for fixed proposals using Gaussians. We have
also added rigorous, numerical error bounds based on t-statistics (from the empirical mean and
variance) asymptotics from [13]. We have added bounds on the equilibrium distribution given bounds
on transition kernel error (we will borrow this from [4]) with the appropriate assumptions.

To reviewers 1, 2, 3, and 4 regarding the Barker function:

You are correct, the logistic function (aka Barker function) is less efficient with \theta' as good
as \theta. However, the tradeoff is that using the Barker function lets us better control our error
rates, and our newest experiments show that we are obtaining orders of magnitude improvements.

To reviewers 1, 2, and 3 regarding the convolution (in)stability:

To make it more stable, we can use L2 regularization and move some of the variance from X_norm to
X_corr. i.e. widen X_corr by convolving with a Gaussian with variance sigma^2, while decreasing the
allowed variance of X_norm by sigma^2. This ensures that X_corr has positive pdf everywhere.  We can
achieve CDF error (X_norm + X_corr compared to X_log) of 1e-7 at sigma = 0.8 or 1e-5 at sigma = 0.9.
We are substantially reworking this part of the paper.

To reviewer 1:

TODO (Daniel: Anything else you want to address)

To reviewer 2:

Regarding the difference between ours and [BDH] sec 6.3, that is close to what we're doing, but they
don't attempt to correct for the difference between normal and logistic distributions with an
additive variable, so they are stuck with a fixed error. That allows us to run our algorithm with
arbitrarily small errors (depending on batch size of course).

TODO (Daniel: Anything else you want to address)

To reviewer 3:

Regarding your Lemma 2 comment, yes, you are correct, the CLT only applies for large n and N, and we
also need n << N. We will clarify this.

To reviewer 7:

Yes increasing the minibatch size is one of the strategies we can use to control variance. The idea
is to increase it just enough, and then to keep it fixed thereafter. We should emphasize this
distinction more, and also to emphasize that this is the least desirable tactic to use (and our
newest results show that we do not need to use it anyway).