### 2016-09-22 nips_2016.tex
More changes to Section 4, I am understanding Johns original intentions better

\documentclass{article}

% if you need to pass options to natbib, use, e.g.: \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
% to avoid loading the natbib package, add option nonatbib: \usepackage[nonatbib]{nips_2016}

\PassOptionsToPackage{numbers, compress}{natbib}
%\usepackage{nips_2016}
% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem}
\usepackage[]{algorithm2e}
\graphicspath{{figures/}}

\lstset{
numbers=left, 
numberstyle=\small, 
numbersep=8pt, 
frame = single, 
language=matlab, 
framexleftmargin=20pt}

\newcommand{\simiid}{\overset{\textrm{i.i.d.}}{\sim}}
\DeclareMathOperator*{\argmin}{arg\,min}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

\title{An Efficient Minibatch Acceptance Test for Metropolis-Hastings}

% The \author macro works with any number of authors. There are two commands used to separate the
% names and addresses of multiple authors: \And and \AND.
% Using \And between authors leaves it to LaTeX to determine where to break the lines. Using \AND
% forces a line break at that point. So, if LaTeX puts 3 of 4 authors names on the first line, and
% the last on the second line, try using \AND instead of \And before the third author name.

%%  David S.~Hippocampus\thanks{Use footnote for providing further
%%    information about author (webpage, alternative
%%    address)---\emph{not} for acknowledging funding agencies.} \\
%%  Department of Computer Science\\
%%  Cranberry-Lemon University\\
%%  Pittsburgh, PA 15213 \\
%%  \texttt{hippo@cs.cranberry-lemon.edu} \\

\author{
  John Canny, Haoyu Chen, Daniel Seita, Xinlei Pan \\
  University of California, Berkeley \\
  \texttt{\{canny,haoyuchen,seita,xinleipan\}@berkeley.edu}
 }

\begin{document}
\maketitle

\begin{abstract}
Markov chain Monte Carlo (MCMC) methods have many applications in
machine learning. We are particularly interested in their application
to modeling very large datasets, where it is impractical to perform
Metropolis-Hastings tests on the full data. Previous work on reducing
the cost of Metropolis-Hastings tests yield variable data consumed per
sample, with only constant factor reductions versus using the full
dataset for each sample.  Here we present a method that can be tuned
to provide arbitrarily small batch sizes, by adjusting either proposal
step size or temperature. Our approach uses the natural noise present
in minibatch likelihood estimates to furnish the randomness in a
Metropolis-Hastings test. Our test uses the noise-tolerant Barker
acceptance test with a novel additive correction variable.  The
resulting test can be combined with minibatch proposals to yield
updates with the same complexity as a simple SGD update. In this paper
we derive the test, analyze its performance, discuss its
implementation, and present several experiments.
\end{abstract}

\section{Introduction}\label{sec:introduction}

Markov chain Monte Carlo (MCMC) sampling is a powerful method for
computation on intractable distributions. We are interested primarily
in large dataset applications, where the goal is to sample a posterior
distribution $p(\theta \mid x_1, \ldots, x_N)$ of parameter $\theta$,
and where the number of data instances $N$ is large.  The
Metropolis-Hastings method (M-H) generates sample candidates from a
proposal distribution $q$ which is in general different from the
target distribution $p$, and decides whether to accept or reject them
based on an acceptance test. The acceptance test is usually a
Metropolis test~\cite{Metropolis1953,hastings70}. Conventional
Metropolis-Hastings requires all $N$ data instances to generate one
posterior sample.

Many state-of-the-art machine learning methods, and deep learning in particular,
are based on minibatch updates (such as SGD) to a model.  Minibatch updates
produce many improvements to the model for each pass over the dataset, and have
high sample efficiency. They also map very well onto hardware such as GPUs. In
contrast, M-H requires calculations over the full dataset to produce a new
sample.  Recent results from~\cite{cutting_mh_2014,icml2014c1_bardenet14}
perform approximate (bounded error) acceptance tests using subsets (minibatches)
of the full dataset. The tests depend on minibatch statistics, and on the value
of an additional uniform random variable $u$. The amount of data consumed for
each test varies significantly from one minibatch to the next, and depends on
the current sample, the proposed sample, and on the random variable $u$. By
contrast,~\cite{conf/uai/MaclaurinA14} performs exact tests but requires a lower
bound on parameter distribution across its domain.  The amount of data reduction
depends on the accuracy of this bound, and such bounds are only available for
relatively simple distributions.

Here we derive a new test which incorporates the variability in minibatch
statistics as {\em a natural part of the test}. Because of this, the amount of
data required for each test is fixed with high probability and in expected
value. {\color{blue} Daniel: we really need to show the amount of data is fixed
with ``high probability'' (the results suggest ``OK probability'').} We use a
Barker test function~\cite{Barker65} rather than a Metropolis test, which makes
our test naturally error tolerant. The idea of using a noise-tolerant test using
Barker's test function was suggested but not explored empirically
in~\cite{Bardenet15} section 6.3. But the asymptotic test statistic CDF and the
Barker function are different, which leads to fixed errors for the approach
in~\cite{Bardenet15}. Here we show that the difference between the distributions
can be corrected with an additive random variable. This leads to a test which is
fast, and whose error can be made arbitrarily small.

Our test is applicable when the variance (over data samples) of the log
acceptance probability is small enough (less than 1). Its not clear at first why
this quantity should be bounded, but we will show that it is ``natural'' for
well-specified models running Metropolis-Hastings sampling with optimal
proposals~\cite{OptimalScaling01} on a full dataset. When we reduce the amount
of data for the test, the variance goes up, and we have to reduce it in one of
several ways. Either:

\begin{itemize}

\item Increase the temperature of the target distribution. Log likelihoods
  scale as $1/T$, and so the variance of the likelihood ratio will
  vary as $1/T^2$. Our model is no longer well-specified (we are doing inference
  at a temperature different from that assumed during data generation), but
  higher temperature can be advantageous for parameter exploration.

\item For continuous probability distributions, reduce the proposal
  step size and variance (for stochastic proposals) compared to an optimal
  proposal. The variance of the log acceptance probability scales as the
  square of proposal step size.

\item Increase the minibatch size. Log acceptance variance scales as $1/k$ vs the
  minibatch size $k$. Increased minibatch size also reduces the error rate for
  the test. {\color{blue} Daniel: I'm not sure if I'd mention this here, because
  before this list, we talk about reducing the data, and then suddenly we're
  talking about increasing it.}

\end{itemize}

Its worth discussing at this point the typical goals of
M-H sampling on very large datasets.  By the Bernstein-von Mises
Theorem, the posterior distribution of the parameter $\theta$ for a
Bayesian inference task is asymptotically normal, and has variance
that scales inversely with the number of data samples $N$. This mode
is extremely sharp for large datasets, which may contain millions or
billions of samples. Simply sampling from this distribution is one
application, but an efficient proposal distribution~\cite{OptimalScaling01} has similar variance to the target
distribution and will diffuse to it extremely slowly from an
initialization value which is (likely to be) many standard deviations
away. If there are any other strong modes, it is very likely for the
sampler to find one of them and become trapped in it when run at the
normal distribution temperature ($T=1$). A common solution is to anneal
the sampler, running first at high temperature (scaling log
likelihoods by $1/T$) which flattens the likelihood landscape.  This is
turn reduces the variance of the log acceptance probability and allows
our acceptance test to be applied.

A second question concerns step size. Once we have fixed temperature,
our variance constraint implies that we have to trade-off proposal
step size $s$ and batch size $b$ ($b \propto p^2$), i.e. we can make
many small steps, or one large step, with a given batch of data. One
of the primary drivers of this work is our belief in the value of
small steps. For applications to neural networks or other models where
the posterior is multimodal, posterior inference is arguably a search
process. Covering the search space densely with small steps is much
more valuable than few sparse steps toward the nearest optimum. In
this mode, Metropolis-Hastings can be used in similar fashion to
Stochastic Gradient Descent. The goal in SGD is to make gradual
progress to a posterior mode with each step, taking small steps so
that the cumulative displacement has progressively lower variance. A
substantial part of the computational work of MCMC on massive datasets
will be similarly in reaching a stationary distribution, which really
means finding a deep posterior mode. Taking noisy small steps will
nevertheless make steady progress to a posterior mode since their bias
is in that direction. We demonstrate this behavior in our logistic regression
experiments.

The contributions of this paper are as follows:

\begin{itemize}
\item We develop a new, more efficient (in samples per test) minibatch acceptance test with quantifiable error bounds. The test
  uses a novel additive correction variable to implement a Barker test based on minibatch mean and variance. 
\item We analyze the test for accuracy and speed.
\item We compared performance of our new test and prior approaches on several datasets. We demonstrate the test is several orders
  of magnitude more efficient than prior work measured as data consumed per test, and that it does not suffer
  from long-tailed minibatch sizes (up to the dataset size). 
\end{itemize}

\section{Preliminaries and Related Work}\label{sec:related_work}

In the Metropolis-Hastings method~\cite{gilks1996markov,brooks2011handbook}, a
difficult-to-compute probability distribution $p(\theta)$ is sampled using a
Markov chain $\theta_1,\ldots,\theta_n$. The sample $\theta_{t+1}$ at time $t+1$
is generated using a candidate $\theta'$ from a (simpler) proposal distribution
$q(\theta'\mid \theta_t)$, filtered by an acceptance test. The acceptance test
is usually a Metropolis test. The Metropolis test has acceptance probability:
\begin{equation}\label{eq:traditional}
    \alpha(\theta_t,\theta') = \frac{p(\theta')q(\theta_t \mid \theta')}{p(\theta_t)q(\theta' \mid \theta_t)} \wedge 1
\end{equation}
where $a \wedge b$ denotes $\min(a,b)$.  With probability
$\alpha(\theta_t,\theta')$, we accept $\theta'$ and set $\theta_{t+1} =
\theta'$, otherwise set $\theta_{t+1}=\theta_t$.  The test is often implemented
with an auxiliary random variable $u \sim \mathcal{U}(0,1)$ with a comparison
$u<\alpha(\theta_t,\theta')$; here, $\mathcal{U}(a,b)$ denotes the uniform
distribution on the interval $[a,b]$.  For notational simplicity, we drop the
subscript $t$ for the current sample $\theta_t$ and denote it as $\theta$.

The acceptance test guarantees detailed balance, which means
$p(\theta)p(\theta'\mid\theta) = p(\theta')p(\theta \mid\theta')$,
where $p(\theta'\mid\theta)$ is the probability of a transition from state
$\theta$ to $\theta'$. Here $p(\theta'\mid\theta) =
q(\theta'\mid\theta)\alpha(\theta,\theta')$ so the detailed balance equation
becomes
\begin{equation}\label{eq:detailed_balance2}
    p(\theta)q(\theta'\mid\theta)\alpha(\theta,\theta') = p(\theta')q(\theta\mid\theta')\alpha(\theta',\theta)
\end{equation}
This condition, together with ergodicity, guarantees that the Markov chain has a
unique stationary distribution $\pi(\theta) = p(\theta)$.

For Bayesian inference, we would like to sample from a parameter distribution
for $\theta$ based on some observed data, so the target distribution is
$p(\theta \mid x_1, \ldots, x_N)$. The acceptance probability is now:
\begin{equation}\label{eq:acceptance_probability}
    \alpha(\theta,\theta') = 
    \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)}
    \wedge 1
\end{equation}
where $p_0(\theta)$ is a prior, and $p(x_i \mid \theta)$ are the probabilities
of the observations. Computing samples this way requires the use of all $N$
training data points, but this is very expensive for large datasets. To address
this challenge,~\cite{cutting_mh_2014,icml2014c1_bardenet14} perform approximate
Metropolis-Hasting tests using sequential hypothesis testing. During each
iteration, they start with a small minibatch of data and test the hypothesis
that the sample $\theta'$ should be accepted based on an approximate version of
the test $u < \alpha(\theta,\theta')$. If the approximate test cannot make a
decision with sufficient confidence, then the minibatch size is increased and
the test repeats. This process continues until a decision. The bounds depend on
either a asymptotic central limit theorem \cite{cutting_mh_2014} or a
concentration bound \cite{icml2014c1_bardenet14}. The latter requires direct
bounds on the log likelihood ratio, which for general distributions requires
knowing $p(x_i \mid \theta)$ and $p(x_i \mid \theta')$ for all $N$ samples. In
addition, both methods suffer the drawback of resolving small log likelihood
ratio differences between the minibatch and full batch versions.  In the worst
case, all $N$ data points may be needed, and we soon show that this worst case
can occur about $\Omega(N)$ times during the performance of $N$ tests.

Following~\cite{icml2014c1_bardenet14}, we write the test $u <
\alpha(\theta,\theta')$ in the equivalent form $\Lambda(\theta,\theta') >
\psi(u,\theta,\theta')$, where\footnote{Our definitions differ from those in
\cite{icml2014c1_bardenet14} by a factor of $N$ to simplify our analysis later.}
\begin{equation}
\Lambda(\theta,\theta') = \sum_{i=1}^N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
\hspace{0.2in} {\rm and} \hspace{0.2in}
\psi(u,\theta,\theta') = \log\left(u\frac{q(\theta'|\theta)p_0(\theta)}{q(\theta|\theta')p_0(\theta')}\right)
\end{equation}
To reduce computational effort, an unbiased estimate of $\Lambda(\theta,\theta')$
based on a minibatch can be used:
\begin{equation}
\Lambda^*(\theta,\theta') = \frac{N}{b}\sum_{i=1}^b \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
\end{equation}
Finally, it will be convenient for our analysis to define the individual
components that contribute to the sums above:
\begin{equation}\label{eq:individual_terms}
\Lambda_i(\theta,\theta') = N \log\left(\frac{p(x_i|\theta')}{p(x_i|\theta)}\right)  
\end{equation}
Thus, $\Lambda(\theta,\theta')$ is the mean of $\Lambda_i(\theta,\theta')$ over
the entire dataset, and $\Lambda^*(\theta,\theta')$ is the mean of
$\Lambda_i(\theta,\theta')$ over its minibatch.

Since minibatches contains randomly selected samples $x_i$, the values
$\Lambda_i$ are independent, identically distributed (iid) random variables.
{\color{blue} Daniel: should we say sampling with or without replacement? Right
now I think the code does it without replacement (but \emph{with} replacement
means it's closer to the CLT, right?} By the Central Limit Theorem, we expect
$\Lambda^*(\theta,\theta')$ to be approximately Gaussian. The acceptance test
then becomes a statistical test of the hypothesis that
$\Lambda(\theta,\theta')>\psi(u,\theta,\theta')$ by establishing that
$\Lambda^*(\theta,\theta')$ is substantially larger than
$\psi(u,\theta,\theta')$.  In~\cite{cutting_mh_2014} an asymptotic central limit
argument was used to derive this gap, while in~\cite{icml2014c1_bardenet14} a
concentration bound was used. In both cases, the resulting tests were shown to
give useful reductions in number of samples required over using the full
dataset, but there were no worst-case bounds given on the average number of
samples needed per iteration.

We next show that for a simple distribution, the lower bound of average
number of data instances consumed for one iteration of~\cite{cutting_mh_2014}
and~\cite{icml2014c1_bardenet14} is $\Omega(N)$, where $N$ is the number of data
points.

\subsection{A Worst-Case Gaussian Example}\label{ssec:gaussian_example}
Consider a Gaussian model where $x_1,\ldots,x_N \simiid \mathcal{N}(\theta,1)$
with a known variance $\sigma^2=1$, true mean $\theta=0.5$, and a uniform prior
on $\theta$. The log likelihood ratio is
\begin{equation}
\Lambda^*(\theta,\theta') =  \frac{N}{b}\sum_{i=1}^b \log\frac{p(x_i|\theta')}{p(x_i|\theta)}=
  %%\frac{1}{2}\left( (x_i-\theta')^2 - (x_i - \theta)^2\right) =
  N(\theta'-\theta)\left(\frac{1}{b}\sum_{i=1}^b x_i-\theta-\frac{\theta'-\theta}{2}\right)
\end{equation}
which is normally distributed over selection of the Normal samples $x_i$.  Since
the $x_i$ have unit variance, their mean has variance $1/b$, and the variance of
$\Lambda^*(\theta,\theta')$ is $\sigma^2(\Lambda^*) = (\theta'-\theta)^2N^2/b$.
In order to pass a hypothesis test that $\Lambda > \psi$, there needs to be a
large enough gap (several $\sigma(\Lambda^*)$) between
$\Lambda^*(\theta',\theta)$ and $\psi(u,\theta',\theta)$. {\color{blue}
Daniel: I changed the variance to be $(\theta'-\theta)^2N^2/b$; before it had
$N$ instead of $N^2$.}

Our goal is to sample from the posterior $p(\theta \mid x_1,\ldots,x_N)$, which
is a normal distribution $\mathcal{N}(\mu, 1/N)$ centered on the sample mean
$\mu$, and with variance $1/N$. In one dimension, an efficient proposal
distribution has the same variance as the target
distribution~\cite{OptimalScaling01}, so we use the proposal
$\mathcal{N}(\theta'-\theta,1/N)$, which is implemented as
$q(\theta'\mid\theta)=\phi((\theta'-\theta)N)$, where $\phi(x)$ is the Normal
density function with zero mean and unit variance. This proposal is symmetric
$q(\theta'\mid\theta)=q(\theta\mid\theta')$, and since we assumed a uniform
prior on $\theta$, we see that $\psi(u,\theta',\theta)$ reduces to $\log u$. Our
worst-case scenario is specified in Lemma~\ref{lem:worst_case}.
{\color{blue} Daniel: this might be a silly question, but why do we assume the
proposal has mean $\theta'-\theta$? I thought it would have a mean of $\theta$
because that is the current sample and the Gaussian is centered at that point?
That will also clarify my confusion about the $\psi$ function.}

\begin{lemma}\label{lem:worst_case}
    For the model in Section~\ref{ssec:gaussian_example}, there exists a fixed
    (independent of $N$) constant $c$ such that with probability $\geq c$ over
    the joint distribution of $(\theta, \theta', u)$, the tests
    from~\cite{cutting_mh_2014,icml2014c1_bardenet14} consume all $N$ samples. 
\end{lemma}

\begin{proof}
See Appendix~\ref{app:worst_case_proof}.
\end{proof}

Similar results can be shown for other distributions and proposals by
identifying regions in product space $(\theta,\theta'-\theta,u)$ such that the
hypothesis test needs to separate nearly-equal values.  It follows that the
accelerated M-H tests in~\cite{cutting_mh_2014,icml2014c1_bardenet14} require at
least a constant fraction $\geq c$ in the amount of data consumed per test
compared to full-dataset tests, so their speed-up is at most $1/c$.
{\color{blue} Daniel: doesn't our method also have this issue? In general, it
seems like any test that requires minibatch information to show deviation from
the full batch version will suffer from this problem. Why not just force us to
stop if we use too much data?}

\subsection{MCMC Posterior Inference}
There is a separate line of MCMC work drawing principles from statistical
physics. By viewing random variables as particles in a system, one can apply
Hamiltonian Monte Carlo (HMC)~\cite{mcmc_hamiltonian_2010} methods which
generate high acceptance \emph{and} distant proposals when run on full batches
of data. Recently Langevin Dynamics~\cite{langevin_2011,conf/icml/AhnBW12} has
been applied to Bayesian estimation on minibatches of data. This simplified
dynamics uses local proposals and avoids MH tests by using small proposal steps
whose acceptance approaches 1 in the limit. However, the constraint on proposal
step size is severe, and the state space exploration reduces to a random walk.
Full minibatch HMC for minibatches was recently described in~\cite{sghmc_2014}
which allows momentum-augmented proposals with larger step sizes. However, step
sizes are still limited by the need to run accurately without MH tests.  Our
work opens the door to applying those methods with much more aggressive step
sizes without loss of accuracy. We demonstrate this in Section ... {\color{blue}
Daniel: do we ever demonstrate this? I guess another issue is that this implies
we can use larger step sizes, but earlier we \emph{also} argue that we need
\emph{small} step sizes (to reduce variance).}

\section{A New Metropolis-Hastings Acceptance Test}\label{sec:our_algorithm}

\subsection{Log-Likelihood Ratios}\label{ssec:log_likelihood_ratios}

For our new M-H test, we denote the exact and approximate log likelihood ratios
as $\Delta$ and $\Delta^*$.  First, $\Delta$ is defined as
\begin{equation}\label{eq:delta1}
    \Delta(\theta,\theta')  =
    \log \frac{p_0(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta \mid
    \theta')}{p_0(\theta)\prod_{i=1}^N p(x_i \mid \theta)q(\theta' \mid\theta)},
\end{equation}
where $p_0, p$, and $q$ match the corresponding functions within
Equation~\ref{eq:acceptance_probability}. We separate out terms dependent and
independent of the data $x$ as:
\begin{equation}\label{eq:delta2}
    \Delta(\theta,\theta') = \sum_{i=1}^N\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') =
    \Lambda(\theta,\theta') - \psi(1,\theta,\theta').
\end{equation}
A minibatch estimator of $\Delta$, denoted as $\Delta^*$, is
\begin{equation}\label{eq:delta3}
    \Delta^*(\theta,\theta') =
\frac{N}{b}\sum_{i=1}^b\log\frac{p(x_i\mid\theta')}{p(x_i\mid\theta)} - \psi(1,\theta,\theta') =
\Lambda^*(\theta,\theta') - \psi(1,\theta,\theta')
\end{equation}
Note that $\Delta$ and $\Delta^*$ are evaluated on the full dataset and a
minibatch of size $b$ respectively. The scaling term $N/b$ ensures that
$\Delta^*(\theta,\theta')$ is an unbiased estimator of $\Delta(\theta,\theta')$.
{\color{blue} Daniel: a brief question repeated from earlier, should we mention
sampling with and/or without replacement here?}

The key to our test is a smooth acceptance function.  We consider functions
other than the classical Metropolis test that satisfy the detailed balance (see
Equation~\ref{eq:detailed_balance2}) condition needed for accurate posterior
estimation. A class of functions leading to detailed balance is specified as
follows:

\begin{lemma}\label{lem:detailed_balance}
    If $g(s)$ is any function such that $g(s) = \exp(s) g(-s)$, then the
    acceptance function $\alpha(\theta,\theta') \triangleq
    g(\Delta(\theta,\theta'))$ satisfies detailed balance.
\end{lemma}

This result is used in~\cite{Barker65} to define the Barker acceptance test.  As
a sanity check, choosing $g(s) = \exp(s) \wedge 1$ --- a function satisfying the
requirement of Lemma~\ref{lem:detailed_balance} --- produces the classical
Metropolis acceptance test $\alpha(\theta,\theta') = g(\Delta(\theta,\theta')) =
\frac{p(\theta')q(\theta \mid \theta')}{p(\theta)q(\theta' \mid \theta)}\wedge
1$. In fact, $g(s) =\exp(s) \wedge 1$ is the optimal acceptance function in
terms of acceptance rate, since it accepts with probability 1 for $\Delta > 0$.

\subsection{Barker (Logistic) Acceptance Function}\label{ssec:barker_function}
For our new MH test we use the Barker logistic~\cite{Barker65} function:
$g(s)=(1+\exp(-s))^{-1}$. Straightforward arithmetic shows that it satisfies the
condition in Lemma~\ref{lem:detailed_balance}.  While it is slightly less
efficient than the Metropolis test when used on the full dataset, we will show
that its smoothness allows it to naturally tolerate substantial variance in its
input argument. This in turn will lead to a much more efficient test on subsets
of data. {\color{blue} Daniel: aside from the experiments, where do we show it
``naturally tolerates substantial variance'' in the input?}

Assume we begin with the current sample $\theta$ and a candidate sample
$\theta'$, and that $V \sim \mathcal{U}(0,1)$ is a uniform random variable. We
accept $\theta'$ if $g(\Delta(\theta,\theta')) > V$, and reject otherwise.
Since $g(s)$ is monotonically increasing, its inverse $g^{-1}(s)$ is
well-defined and unique. So an equivalent test is to accept $\theta'$ iff
\begin{equation}\label{eq:equivalent_test}
    \Delta(\theta,\theta') > X = g^{-1}(V)
\end{equation}
where $X$ is a random variable with the logistic distribution (its CDF is the
logistic function). To see this notice that $\frac{dV}{dX} = g'$, that $g'$ is
the density corresponding to a logistic CDF, and finally that $\frac{dV}{dX}$ is
the density of $X$. The density of $X$ is symmetric, so we can equivalently test
whether
\begin{equation}\label{eq:the_exact_test}
    \Delta(\theta,\theta') + X > 0
\end{equation}
for a logistic random variable $X$. {\color{blue} Daniel: I think the text
before Equation~\ref{eq:equivalent_test} should be similar to what we had in
our NIPS submission because I find it much clearer.}

\subsection{A Minibatch Acceptance Test}\label{ssec:deltas_minibatch}

We now describe acceptance testing using the minibatch estimator
$\Delta^*(\theta,\theta')$. From Equation~\ref{eq:delta3},
$\Delta^*(\theta,\theta')$ can be represented as a constant term plus the mean
of $b$ IID terms $\Lambda_i(\theta,\theta')$ of the form
$N\log\frac{p(x_i|\theta')}{p(x_i|\theta)}$. As $b$ increases,
$\Delta^*(\theta,\theta')$ therefore has a distribution which approaches a
normal distribution by the Central Limit Theorem. We now describe this using an
asymptotic argument and defer specific bounds between the CDFs of
$\Delta^*(\theta,\theta')$ and a Gaussian to Section~\ref{sec:analysis}.

In the limit, since $\Delta^*$ is normally distributed about its mean $\Delta$,
we can write
\begin{equation}\label{eq:relationship}
    \Delta^* = \Delta + X_{\rm norm}, \quad X_{\rm norm} \sim \mathcal{\bar{N}}(0, \sigma^2(\Delta^*)),
\end{equation}
where $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ denotes a distribution which is
approximately normal with variance $\sigma^2(\Delta^*)$.  But to perform the
test in Equation~\ref{eq:the_exact_test} we want $\Delta + X$ for a logistic
random variable $X$ (call it $X_{\rm log}$ from now on). In~\cite{Bardenet15} it
was proposed to use $\Delta^*$ in a Barker test anyway and tolerate the fixed
error caused by this approximation.

Our approach is to instead decompose $X_{\rm log}$ as
\begin{equation}\label{eq:deconvolution}
    X_{\rm log} = X_{\rm norm}+X_{\rm corr},
\end{equation}
where we assume $X_{\rm norm} \sim \mathcal{N}(0, \sigma^2)$ and $X_{\rm corr}$
is a zero-mean ``correction'' variable with density $C_{\sigma}(X)$.  The two
variables are added (i.e., their distributions convolve) to form $X_{\rm log}$.
This decomposition requires an appropriate $X_{\rm corr}$ distribution. We show
in Section~\ref{sec:correction} that it is possible to use deconvolution to
derive a highly accurate representation of the density $C_\sigma(X)$.
Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
$X_{\rm corr}$ distributions that sum to $X_{\rm log}$.

Using $X_{\rm corr}$ samples from $C_\sigma(X)$, the acceptance test is now
\begin{equation}\label{eq:criteria}
    \Delta + X_{\rm log} = (\Delta + X_{\rm norm}) + X_{\rm corr} = \Delta^* + X_{\rm corr} >0.
\end{equation}
Therefore, assuming the variance of $\Delta^*$ is small enough, if we have an
estimate of $\Delta^*$ from the current data minibatch, we test acceptance by
adding a random variable $X_{\rm corr}$ and then accept $\theta'$ if the result
is positive (and reject otherwise).

If $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ is exactly $\mathcal{N}(0,
\sigma^2(\Delta^*))$, the above test is exact as well, and as we show in
Section~\ref{sec:analysis}, if there is a maximum error $\epsilon$ between the
CDF of $\mathcal{\bar{N}}(0, \sigma^2(\Delta^*))$ and the CDF of $\mathcal{N}(0,
\sigma^2(\Delta^*))$, then the acceptance test has an error of at most
$\epsilon$ relative to the full batch version. {\color{blue} Daniel: this
subsection is improved from earlier. My main concern now is that $X_{\rm norm}$
is described as being both approximately Gaussian $\bar{\mathcal{N}}$ and
Gaussian $\mathcal{N}$ here. We may want to clear that up; in our NIPS
submission we actually used a new variable $\varepsilon$ for that and assigned
it to be the $\mathcal{N}$ distribution.}

\begin{figure}[t]
    \centering
    \includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
    \caption{
    Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
    convolve to form the standard logistic distribution. We use three standard
    deviation values of $X_{\rm norm}$. The two red curves convolve to form the
    logistic, etc. The $y$-axis is capped at 0.5 for readability. This figure
    must be viewed in color. {\color{blue} Daniel: TODO Change figure to use
    LaTeX for the titles.}
    }
    \label{fig:deconvolution}
    %\vspace{-10pt}
\end{figure}

\section{Computing the Correction Distribution}\label{sec:correction}

Our proposed test in Equation~\ref{eq:criteria} requires knowing the
distribution of the correction variable $X_{\rm corr}$ such that $X_{\rm norm} +
X_{\rm corr} = X_{\rm log}$, where $X_{\rm norm} \sim \mathcal{N}(0,\sigma^2)$
and $X_{\rm log}$ has a standard logistic CDF, $(1+\exp(-X))^{-1}$. In
Section~\ref{sec:analysis}, we show that the accuracy of the test depends on the
absolute error between the CDFs of $X_{\rm norm} + X_{\rm corr}$ and $X_{\rm
log}$. Consequently, we need to minimize this in our construction of $X_{\rm
corr}$. More formally, let $\Phi_\sigma(X) = \Phi(X/\sigma)$ where $\Phi$ is the
standard normal CDF\footnote{Hence, $\Phi_\sigma$ is the CDF of a zero-mean
Gaussian with variance $\sigma$.}, $S(X)$ be the logistic function, and
$C_{\sigma}(X)$ be the \emph{density} of the correction $X_{\rm corr}$
distribution. Our goal, based on Equation~\ref{eq:deconvolution}, is to solve
the following optimization problem:
\begin{equation}\label{eq:overall_corr_problem}
    C_\sigma^* = \argmin_{C_\sigma} |\Phi_{\sigma} * C_{\sigma} - S|
\end{equation}
where $*$ denotes convolution. {\color{blue} Daniel: This is very unusual use of
convolutions (i.e., CDF * DENSITY = CDF). It's valid but I was very surprised to
know that.} For computation of $C_\sigma$, we assume that its input $Y$ and
another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
{\color{blue} Daniel: I am a little confused at the different interval sizes.}
In continuous form, the optimization problem is therefore
\begin{equation}\label{eq:conv1}
    C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|
\end{equation}
To get $C_\sigma^*$ in a practical form, we discretize $X$ and $Y$ into $4N+1$
and $2N+1$ values respectively. Writing $X_i = i*V/N$ and $Y_j = j*V/N$, the
condition in Equation~\ref{eq:conv1} can be re-expressed as:
\begin{equation}\label{eq:conv2}
    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
\end{equation}
Defining a matrix $M$ and vectors $u$ and $v$ such that $M_{ij} =
\Phi_{\sigma}(X_i-Y_j)$, $u_j = C_{\sigma}(Y_j)$, and $v_i = S(X_i)$,
Equation~\ref{eq:conv2} is equivalent to minimizing the $L_1$ norm:
\begin{equation}\label{eq:optimization_l1}
    u^* = \argmin_{u} \|Mu-v\|_1.
\end{equation}
{\color{blue} Daniel: Equation~\ref{eq:optimization_l1} is finding
the $C_\sigma$ minimizing the $L_1$ norm, but the previous one,
Equation~\ref{eq:conv2}, is finding the $C_\sigma$ that is minimizing the
$L_\infty$ norm, i.e., the maximum value! If we want this to be the $L_1$ norm
(and I think we do) then Equation~\ref{eq:conv2} needs an outer sum over the $i$
terms. In addition, Equation~\ref{eq:conv1} also needs to be modified, perhaps
eliminate the ``sup'' but keep $x \in [-2V,2V]$}

We have the additional constraint that $u_j > 0$ for all $j$, since $u$
represents a density. We first explored optimizing this problem with linear
programming to find a $u$ minimizing $|Mu - v|$ subject to $u \ge 0$. This was
tractable up to a few hundred dimensions for $u$. However, the discretization
error is bounded by the curvature of the underlying functions which are here
slightly less than one, i.e.  the errors are of the order of $(V/N)^2$.
{\color{blue} Daniel: unfortunately, I do not understand what this means (i.e.,
the curvature and how the $(V/N)^2$ is obtained).} Here we chose $V=20$ to
provide adequate containment of the distributions (the CDFs are extremely close
to either zero or one outside this range). So with 200 points, we have a
discretization error of approximately 0.01, which is relatively large.  To yield
higher resolution and lower error, we switched to a least squares solution:
\begin{equation}\label{eq:optimization_l2}
    u^* = \arg_u \min \; \|Mu-v\|_2^2 + \lambda \|u\|_2^2,
\end{equation}
where we add the standard regularization factor $\lambda$ for stability in high
dimensions. The solution is well-known from the normal equations: $u^* = (M^TM +
\lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_1$ norm for
Equation~\ref{eq:optimization_l1}.

With this approach, there is no guarantee that $u^* \geq 0$. However, we have
some flexibility in the choice of $\sigma$ in Equation~\ref{eq:conv1}.
{\color{blue} Daniel: is this the correct equation?} As we decrease the variance
of $X_{\rm norm}$, the variance of $X_{\rm corr}$ grows by the same amount and
is in fact the result of convolution with a Gaussian whose variance is the
difference.  Thus as $\sigma$ decreases, $C(X)$ grows and approaches the
derivative of a logistic function at $\sigma = 0$. It retains some very weak
negative values for $\sigma > 0$ but removal of those values leads to very small
error.

\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\begin{algorithm}[t]
\Input{Number of samples $T$, minibatch size $m$, error bound $\delta$, pre-computed correction $C_1(X)$ distribution, initial sample $\theta_1$.}
\Output{A chain of $T$ samples $\{\theta_1, \ldots, \theta_T\}$ from $p(\theta)$\;}
\For{$t=\{1, \ldots, T\}$}{
    Propose a candidate $\theta'$\ from proposal distribution $q(\theta'\mid\theta_t)$\;
    Draw a minibatch of $m$ points $x_i$, compute $\Delta^*(\theta_t,\theta')$ and sample variance $\sigma^2(\Delta^*)$\;
    Estimate moments $E|\Lambda_i-\Lambda|$ and $E|\Lambda_i-\Lambda|^3$ from the sample, and error $\epsilon$ from equation (\ref{eq:clt-bounds3})\;
    \While {$\sigma^2(\Delta^*) \ge 1$ {\bf or} $\epsilon > \delta$}{
    	Draw $m$ more samples to augment the minibatch, update $\Delta^*$, $\sigma^2(\Delta^*)$ and $\epsilon$ estimates\;
    }
    Draw $X_{\rm norm} \sim \mathcal{N}(0,1-\sigma^2(\Delta*))$ and
    $X_{\rm corr}$ from the correction distribution $C_1(X)$\;

        \eIf{$\Delta^* + X_{\rm norm} + X_{\rm corr} > 0$}{
            Accept the candidate, $\theta_{t+1} = \theta'$\;
        }{
            Reject and re-use the old sample, $\theta_{t+1} = \theta_t$\;
        }

}
\caption{A description of M-H sampling with our acceptance test. {\color{blue}
Daniel: I think we may want to be careful about the use of $C_1(X)$ because that
implies that we have a fixed $X_{\rm corr}$ density, but that $C_1(X)$ will
change depending on the iteration and variance of $\Delta^*$, right? Hopefully
not by much, but it still changes sometimes?}}
\label{alg:our_algorithm}
\end{algorithm}

\begin{table}[t]
    \caption{\color{blue} Daniel: TODO need a good title.}
    \label{tab:xcorr}
    \centering
    \begin{tabular}{l l l l}
    \toprule
    $N$ & $\sigma$ & $\lambda$ & $L_1$ error  \\
    \midrule
    4000 & 0.9 & 1 & 1.0e-4  \\
    4000 & 0.8 & 0.03 & 5.0e-6 \\
    \bottomrule
    \end{tabular}
\end{table}

{\color{blue} Daniel: TODO improve/describe Table~\ref{tab:xcorr} and try to
relocate Algorithm~\ref{alg:our_algorithm} to be somewhere nicer.}

\section{Analysis}\label{sec:analysis}

We now derive error bounds for our M-H test, and for the approximate target
distribution that it generates. From Section~\ref{sec:correction}, we know that
it is possible to generate the correction samples $X_{\rm corr}$ with a CDF
error of at most single-precision floating point error, and possibly better. We
therefore treat the $X_{\rm corr}$ variable as a sample from the exact
correction distribution and we will not analyze its errors.

In the most similar prior works, \cite{cutting_mh_2014} uses asymptotic arguments based
on the central limit theorem to argue that its approximate acceptance test error tends to
zero as batch size increases. But no quantitative bounds are given. In
\cite{icml2014c1_bardenet14}, explicit bounds are given, but they depend on a bound on
the quantity:
\begin{equation}
  \max_{1\leq i\leq N}|\log p(x_i|\theta') - \log p(x_i|\theta)|
\end{equation}
while such bounds can be derived easily for simple models, its unclear how to derive them
for a complex model such as a neural network.

In this paper, we rely on quantitave forms of the central limit theorem which rely on measurable
statistics from the minibatch $\{x_i\}$. Thus a sampler using our approach does not need to ``see''
data beyond the current minibatch. This supports use of the sampler on very large datasets, and
in online mode where the dataset is presented as a stream.

The outline of this section is as follows: we first present bounds on the absolute
and relative error (in CDF) of the distribution of $\Delta^*$ vs a Normal distribution. We then show that
these errors bounds are preserved after the addition of other random variables, in particular
$X_{\rm norm}$ and $X_{\rm corr}$. From this it follows that the acceptance test has the same error bound.

The quantitative central-limit result below comes from \cite{explicit-clt05}:

\begin{lemma}\label{lem:quant_clt}
Let $X_1,\ldots,X_n$ be
a set of zero-mean, independent, identically-distributed random variables with sample mean $\hat{X}$ and
variance $\sigma^2(\hat{X})$ where:
\begin{equation}
    \hat{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad \sigma(\hat{X}) = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\hat{X})^2\right)^{\frac{1}{2}}
\end{equation}
then $t=\hat{X}/\sigma(\hat{X})$ has an approximate Student's distribution which approaches a normal
distribution in the limit. Then from~\cite{explicit-clt05}:
\begin{equation}\label{eq:clt-bounds}
    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
\end{equation}
\end{lemma}

So as long as we know the first and third absolute moments $E|X|$ and $E|X|^3$, we can bound the
error of the normal approximation, which decays as $O(n^{-\frac{1}{2}})$. Making the change of variables
$y = x \sigma(\hat{X})$, Equation~\ref{eq:clt-bounds} becomes
\begin{equation}\label{eq:clt-bounds2}
   \sup_y\left|{\rm Pr}(\hat{X}< y) - \Phi\left(\frac{y}{\sigma(\hat{X})}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}
\end{equation}
which shows that the distribution of $\hat{X}$ approaches the normal distribution $\mathcal{N}(0,\sigma^2(\hat{X}))$
whose variance is $\sigma^2(\hat{X})$, measured from the sample.

To apply this to our test, let $X_i = \Lambda_i(\theta,\theta') - \Lambda(\theta,\theta')$,
then the $X_i$ are zero-mean, IID variables.
Take $n=b$ the minibatch size, and then:
\begin{equation}
    \hat{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta') 
\end{equation}
so that $\sigma(\hat{X}) = \sigma(\Delta^*)$.
\begin{corollary}
We can now substitute into Equation~\ref{eq:clt-bounds2} and displace by the mean, giving:
\begin{equation}\label{eq:clt-bounds3}
    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)\right| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} = \epsilon(\theta,\theta',b)
\end{equation}
\end{corollary}

We have shown that the distribution of $\Delta^*$ approximates a Normal
distribution with mean $\Delta$ and variance $\sigma^2(\Delta^*)$, and bounded
its error with estimable quantities. Both $E|X|$ and $E|X|^3$ can be estimated
as means of $|\Lambda_i - \Lambda|$ and $|\Lambda_i - \Lambda|^3$ respectively
on each minibatch. We expect this will often be accurate enough on minibatches
with hundreds or thousands of points, but otherwise bootstrap CIs can be
computed from those sequences. Since the bounds are monotone in $E|X|$ and
$E|X|^3$, using upper bootstrap CI limits will provide high-confidence error
bounds.

{\color{blue} Daniel: TODO provide a better transition for this paragraph.}

\begin{lemma}\label{lem:cdf_bounds}
Let $P(x)$ and $Q(x)$ be two cumulative distributions satisfying
$\sup_x|P(x)-Q(x)|\leq \epsilon$ with $x$ in some real range. Let $R(y)$ be the
{\em density} of another random variable $y$. Let $P'$ be the convolution $P*R$
and $Q'$ be the convolution $Q*R$. Then $P'(z)$ (resp. $Q'(z)$) is the CDF of
sum $z=x+y$ of independent random variables $x$ with CDF $P(x)$ (resp. $Q(x)$)
and y with density $R(y)$.  Then
\begin{equation}
    \sup_x|P'(x)-Q'(x)|\leq \epsilon
\end{equation}
\end{lemma}

\begin{proof}
We first observe that
\begin{equation}\label{eq:pdf_difference}
    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
\end{equation}
and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
\begin{equation}
-\epsilon = \int_{-\infty}^{+\infty} -\epsilon R(x) dx \leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon
\end{equation}
\end{proof}

\begin{corollary}
If $\sup_y|{\rm Pr}(\Delta^* < y) - \Phi(\frac{y-\Delta}{\sigma(\Delta^*)})|
\leq \epsilon(\theta,\theta',b)$, then
\begin{equation}
    \sup_y|{\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) - S(y-\Delta)| \leq \epsilon(\theta,\theta',b)
\end{equation}
where $S(x)$ is the standard logistic function, and $X_{\rm norm}$ and $X_{\rm corr}$ are generated as per Algorithm 1. 
\end{corollary}

\begin{proof}
We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
\begin{equation}
    P(y) = Pr(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
\end{equation}
and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
where $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of
$P$ and $Q$
\begin{equation}
    P'(y) = Pr(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
\end{equation}
Now we convolve with the distribution of $X_{\rm corr}$ which gives:
\begin{equation}
    P''(y) = Pr(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
\end{equation}
Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
norm}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
log} > 0)$ by at most $\epsilon$.
\end{proof}

In fact we can do better than this approximation (showing the error decreases as $O(n^{-1})$) by using a more
precise limit distribution under an additional assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$
denote the $i^{th}$ absolute moment of $X$. If Cramer's condition holds, i.e. if
\begin{equation}
  \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1.
\end{equation}

Equation 2.2 in Bentkus et al.'s work on Edgeworth expansions~\cite{Bentkus97}
provides the following:

\begin{lemma}\label{lem:clt_edgeworth}
Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
random variables with sample mean $\hat{X}$ and with $t$ defined as in Lemma 3.
If $X$ satisfies Cramer's condition, then
\begin{equation}\label{eq:clt-bounds_edgeworth}
    \sup_x\left|{\rm Pr}(t<x) - G\left(x, \frac{\mu_3}{b_2^{3/2}}\right)\right| \leq \frac{c(\epsilon,b_2,b_3,b_4,b_{4+\epsilon})}{n}
\end{equation}
where
\begin{equation}
    G_n(x,y) = \Phi(x) + \frac{y(2x^2+1)}{6\sqrt{n}}\Phi'(x).
\end{equation}
\end{lemma}

Lemma~\ref{lem:clt_edgeworth} shows that the average of the $X_i$ has a more
precise, skewed CDF limit $G_n(x,y)$ where the skew term has weight proportional
to a certain measure of skew derived from the moments:
$\frac{\mu_3}{b_2^{3/2}}$. Note that if the $X_i$ are symmetric, the weight of
the correction term is zero, and the CDF of the average of the $X_i$ converges
to $\Phi(x)$ at a rate of $O(n^{-1})$.

Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
$n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where
$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
\begin{equation}\label{eq:GNderivatives}
    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
\end{equation}
From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
$\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In Algorithm 1, we
correct for the difference in $\sigma$ between $\Delta^*$ and the variance
needed by $X_{\rm corr}$ using $X_{\rm norm}$. This same method works when we
wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
component functions of $G_n(x,y)$ are derivatives of a (unit variance)
$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm norm}$ as
per Algorithm 1 preserving the limit in Equation~\ref{eq:GNderivatives}.

The deconvolution approach can be used to construct a correction variable
$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
saved offline. As we have shown above, errors in the ``limit'' function
propagate directly through as errors in the acceptance test.  To achieve a test
error of say $1e-6$ (close to single floating point precision), we need a $y'$
spacing of $1e-6$. It should not be necessary to tabulate values all the way to
$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
resolution is 10,000, this leads to a table with about 1 billion values, which
can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
in $y$, it should be possible to achieve similar accuracy with a much smaller
table. We leave further analysis and experiments with $g(x,y)$ as future work.

%% \textbf{The Convolution}. Figure~\ref{fig:deconvolution} demonstrates
%% Equation~\ref{eq:deconvolution} and provides three (color-coded) examples of densities for $X_{\rm
%% norm}$ and $X_{\rm corr}$. The density of $X$, a logistic random variable with mean $\mu = 0$ and
%% scale $s=1$, is known, as are the $X_{\rm norm}$ densities for different standard deviations. The
%% deconvolution provides us with the $X_{\rm corr}$ distributions. Note, however, that as the standard
%% deviation of $X_{\rm norm}$ increases, the $X_{\rm corr}$ distribution becomes increasingly unstable
%% and ``bumpier'', because the logistic curve has fatter tails than Gaussians. In addition, we cap
%% the standard deviation possibilities of $X_{\rm norm}$ at $\approx 1.2$ in our experiments. In our
%% experiments, we discretize the possible standard deviations of $X_{\rm norm}$, and we also limit the
%% densities to consider the range $[-10,10]$ as shown in Figure~\ref{fig:deconvolution}.

%% \textbf{Variance Preconditioning}. Returning to the discussion from Section~\ref{sec:introduction},
%% our test requires a variance check. Equation~\ref{eq:deconvolution} means that the variance of $X$,
%% which is $\pi^2/3\approx 3.29$ is an upper bound on the variances of $X_{\rm norm}$. Therefore, the
%% test cannot work if the variance of $X_{\rm norm}$ is too large. The standard deviation of $X$
%% is about $\sqrt{3.29}\approx 1.81$, so this bounds ${\rm std}(X_{\rm norm})$ (as well as ${\rm
%% std}(\Delta')$). From the discussion above, we bound the standard deviation at 1.2 so that $X_{\rm
%% corr}$ can handle the remaining noise. Recall from Section~\ref{ssec:deltas_minibatch} that we
%% estimate the ${\rm std}(X_{\rm norm})$ each iteration. If the variance/standard deviation test
%% fails, then we can (a) skip the iteration, (b) increase the minibatch size, (c) change the proposal
%% to decrease step sizes, or (d) increase temperature, as we now discuss.

%% \textbf{Temperature}. One way to satisfy the variance precondition is to increase the temperature of
%% the target distribution. For a temperature $T>1$, the augmented target distribution would become
%% \begin{equation}\label{eq:log_temperature}
%% \log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
%% \end{equation}
%% with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
%% smaller in absolute value, thus reducing variance. The flatter posterior results in easier mixing.

\section{Experiments}\label{sec:experiments}

We conduct two sets of experiments to explore the benefits of our minibatch MH test and to
benchmark it with previous work. In Section~\ref{ssec:gaussians}, we show that our test enables
samples to converge to the posterior distribution of a heated Gaussian mixture model. In
Section~\ref{ssec:logistic}, we analyze its efficiency on logistic regression. 
%In %Section~\ref{ssec:nets}, we apply the test for deep learning. 
Appendices~\ref{app:gaussian},~\ref{app:logistic} %and~\ref{app:nnet} 
contain more detailed information on these respective experiments.

\subsection{Mixture of Gaussians}\label{ssec:gaussians}

We start with a simple Gaussian mixture model, borrowing an experiment from~\cite{langevin_2011}.
The parameter is 2-D, $\theta = (\theta_1,\theta_2)$, and the parameter/data generation process is
\begin{equation}\label{eq:data_generation}
(\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
\end{equation}
We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta = (0,1)$, we draw 1,000,000
data points so that the target distribution is $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with
the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This
results in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we apply a
temperature $T=10,000$ to reduce ${\rm std}(\Delta')$. Taking logs, we get the target as shown in the
far left of Figure~\ref{fig:gauss_mix_1}.

We run MCMC with our MH test using minibatch size $m=100$. %We also run this using standard
%full-batch MCMC (``standard MCMC'') with the standard test from Equation~\ref{eq:traditional}, 
We also run this using the method from~\cite{cutting_mh_2014} (``Cut MH'')  and the method from 
~\cite{icml2014c1_bardenet14}. For the method in \cite{cutting_mh_2014}, we use $m=100$ and
increment minibatches by that amount within a test if necessary. As for the method in ~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the minibatch size geometrically with a ratio of $\gamma = 1.5$. The tolerance for making a
decision in \cite{cutting_mh_2014} is $\epsilon=0.005$, and the error bound control parameter in ~\cite{icml2014c1_bardenet14} is $p = 2, \delta = 0.01$. To make comparisons easier, all three use the same random walk proposer with covariance $\Sigma={\rm diag}(0.3, 0.3)$. This is a poor proposer, but it is sufficient for
our purposes as the quality of the samples will be primarily due to the MH test. All methods are run
5000 times to collect 5000 samples.

\begin{figure}[t]
    \centering
    % \includegraphics[width=1\linewidth]{cloud_v01.png}
    \includegraphics[width=1\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
    \caption{
    The log posterior contours and scatter plots of sampled $\theta$ values using different methods. 
    }
    \label{fig:gauss_mix_1}
    %\vspace{-10pt}
\end{figure}

\begin{figure}[t]
    \centering
   % \includegraphics[width=1\linewidth]{adaptive_and_ours_information_v01.png}
   \includegraphics[width=1\linewidth]{GaussianMixtureResult/minibatch_size_gaussian.png}
    \caption{
    Histograms of minibatch sizes %and ${\rm std}(\Delta')$ values, along with ``${\rm std}(\Delta')$
   % violations'' (right).
    }
    \label{fig:gauss_mix_2}
    %\vspace{-10pt}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
	\caption{Value of first and second part of equation (9) in ~\cite{icml2014c1_bardenet14} with respect to actual minibatch size}
	\label{fig:gauss_bardenet}
\end{figure}

%\begin{figure}[t]
    %\centering
    %\includegraphics[width=1\linewidth]{CI_v01.png}
   % \caption{
   % Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
    %}
   % \label{fig:gauss_mix_3}
    %\vspace{-10pt}
%\end{figure}

Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$ samples for the four
methods, with darker regions indicating a greater density of points. The three methods obtain the
same rough form of the posterior, so our MH test can indeed result in the same posterior as the
other two methods. %However, the fourth method (i.e. just treat the $\Delta'$ which comes from minibatch as $\Delta$ which comes from the complete data) reveals wider spread distribution than the posterior. From Figure~\ref{fig:gauss_mix_1} we illustrate our proposed method can obtain the samples which have almost the same distribution as the standard MH test from the complete data, adaptive minibatch test, and the true posterior. Meanwhile, we show the benefits of our minibatch method compared with the naive minibatch method.  
Actually, based on a measure of divergence from the ground-truth distribution, our method is better than the other three methods in terms of this measurement. The measurement is to calculate the poisson likelihood of the MCMC-generated parameters over a parameter range, with $\theta_1 \in (-1.5, 2.5)$ and $\theta_2 \in (-2.5, 2.5)$. The log likelihood value for our method, cut MH method and adaptive MH method are -2144.6, -4007.4, -2598.4, respectively.

Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used by
the three methods in each iteration. It is obvious that our methods consumes significantly less data, and most minibatch sizes are smaller than 1000, with an average minibatch size of 420. The other two methods occasionally consumes a large proportion of the entire data set, and the average minibatch sizes are 16378, 16755 for cut MH method, adaptive MH method, respectively. In terms of average minibatch size, our method is nearly 39 times faster the cut MH method, and nearly 40 times faster than the adaptive MH method.

It is useful to notice that, in~\cite{icml2014c1_bardenet14}, they need to evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set at every iteration in order to calculate the error bound in equation (9) in ~\cite{icml2014c1_bardenet14}. In Figure ~\ref{fig:gauss_bardenet}, the value of the first part and second part of equation (9) in ~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual minibatch data used. The second part which contains  the $C_{\theta,\theta'}$ term is larger than the first part, which means that the second part is non-negligible and has to be calculated at every iteration. This is one of the very time-consuming parts in this method.

%Usually, the method can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting in lots of wasted computation.  For this particular experiment, that happened eight times.
%The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
%\ge 1.2$ for only consuming one minibatch data. For six minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
% our MH test five times and averaged the number of ``violations.'' As the size increases, the
% number of violations decreases, as expected.

\iffalse
Figure~\ref{fig:gauss_mix_3} (to the left) shows a histogram of ${\rm
  std}(\Delta')$. Figure~\ref{fig:gauss_mix_3} (to the middle) shows
the confidence interval of our estimated ${\rm std}(\Delta')$ at each
iteration \footnote{In order to show the confidence interval clearly,
  we use moving smooth to smooth the confidence interval, and plot
  each value by interval of 50 iterations}. The blue line represents
the standard deviation comes from the complete data, which can be
treated as the ``true'' ${\rm std}(\Delta')$, the green line
represents our estimation of ${\rm std}(\Delta')$ from the minibatch,
and the two red dash lines represent the confidence interval of ${\rm
  std}(\Delta')$ by adding / subtracting one ${\rm
  std(std(}\Delta'))$. This figure indicates that our estimation of
${\rm std}(\Delta')$ from one minibatch data is
accurate. Figure~\ref{fig:gauss_mix_3} (to the right) shows the value
of $\rho/\sigma^3$ at each iteration. 
\fi

\subsection{Logistic Regression}\label{ssec:logistic}

We next use logistic regression for the binary classification of 1s versus 7s in the MNIST
dataset~\cite{lecun-mnisthandwrittendigit-2010}. The data has 12007 and 1000 training and testing
points, respectively (we used a random subset of the test data). The proposer is again a random walk
with covariance matrix $0.05I$ for the $784\times 784$ identity matrix $I$. We set the posterior
temperature at $T=1000$. We set the minibatch size $m=100$ and compare with ~\cite{cutting_mh_2014} with
tolerance 0.005 and with ~\cite{icml2014c1_bardenet14} with error bound control parameter $p = 2, \delta = 0.01$.

Figure~\ref{fig:logistic_fig} shows the prediction accuracy and log likelihood on the test set as a
function of the cumulative training data points processed. Our test increments the cumulative data
by a fixed amount per iteration, but the other two methods may require more data per iteration.  We
see that our minibatch MH test is more efficient; it has similar or better performance while
consuming fewer data points.

\begin{figure}[t]
    \centering
    \begin{subfigure}{.4\textwidth}
     \centering
     \includegraphics[width=1\linewidth]{LogisticRegressionResult/acc_logistic}
    \end{subfigure}
    \begin{subfigure}{.4\textwidth}
     \centering
    \includegraphics[width=1\linewidth]{LogisticRegressionResult/ll_logistic}
\end{subfigure}
\caption{Logistic regression performance (accuracy/log likelihood) and minibatch size analysis}
\label{fig:logistic_fig}
\vspace{-10pt}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=1\linewidth]{LogisticRegressionResult/minibatch_size_logistic.png}
	\caption{Histograms of minibatch sizes of logistic regression parameter estimation}
	\label{fig:logistic_minibatch}
\end{figure}

%\begin{figure}[t]
  %  \centering
    %\includegraphics[width=1\linewidth]{CI_v02.png}
    %\caption{
   % Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
    %}
    %\label{fig:exp_3}
    %\vspace{-10pt}
%\end{figure}

The plot of Figure ~\ref{fig:logistic_minibatch} shows the histogram of minibatch sizes. Choosing an initial minibatch size of 100, 
the cut MH method, adaptive MH method achieves final average minibatch size of 585 and 2731 respectively, while our method achieves 
that of 100, showing significant better performance than the benchmark test methods.

%The third plot of Figure~\ref{fig:logistic_minibatch} shows performance based on minibatch size. The
%parameter is evaluated on the test set after training iterations 700, 1000, and 1500. For those
%three cases, when the minibatch size increases beyond about 10, the performance of the sampled data
%will not improve with a large minibatch size. This makes sense because the error of our MH test is
%proportional to our estimation error of ${\rm std}(\Delta')$, and as the minibatch increases, ${\rm
%std}(\Delta')$ decreases. Beyond a certain standard deviation (or minibatch) threshold, the error in
%our minibatch MH test is negligible.

%Similar as the Subsection~\ref{ssec:gaussians}, we use Figure~\ref{fig:exp_3}
%to reveal the accuracy of our estimation of ${\rm std}(\Delta')$ and stability
%of $\rho/\sigma^3$.

%% \subsection{Neural Network Optimization}\label{ssec:nets}
%% 
%% In this experiment, we apply our minibatch MH test with a Stochastic Gradient Hamiltonian
%% Monte Carlo (SGHMC)~\cite{sghmc_2014} proposer to sample the complete parameters for a fully
%% connected neural network. We use the Higgs data set from the UCI dataset~\cite{Lichman:2013}, which
%% is a binary classification task with one million instances, each of which are 28-dimensional. The
%% network we use has four layers: the first is the input, the last is the softmax, and the
%% intermediate ones apply the sigmoid activation unit.  We also employ dropout and batch
%% normalization~\cite{icml2015_ioffe15}.  In order to control ${\rm std}(\Delta')$, we initialize the
%% temperature at 1000, and adjust it at iteration $i$ according to $T_i = \max\{1,
%% 1000/(i+1)^{0.5}\}$.
%% 
%% For our comparison baseline, we train the same network architecture using the Adaptive Gradient
%% Descent method~\cite{adapGrad}. Since the frequencies of the features are quite different, we use
%% the same strategy as in Adaptive Gradient to rescale the gradient before applying SGHMC. We
%% implement our MH test and SGHMCin the open-source BIDMach project~\cite{canny2013bidmach}.  
%% 
%% Figure~\ref{fig:nnet_fig} illustrates our experiment results and reveals that SGHMC with our
%% minibatch MH test achieves higher log likelihood than the baseline. By comparing SGHMC with and
%% without our minibatch MH test, we see that our MH test has limited benefit. Because we decrease the
%% step size quickly, the acceptance rate of SGHMC is already high, so there is little need to test.
%% % Daniel's note: Haoyu said he cannot increase the step size without it violating the variance
%% % condition.
%% 
%% \begin{figure}[t]
%%     \centering
%%     \includegraphics[width=1\linewidth]{exp3}
%%     \caption{Log likelihood versus number of passes for optimization of neural network.}
%%     \label{fig:nnet_fig}
%%     \vspace{-10pt}
%% \end{figure}

\section{Conclusions}\label{sec:conclusion}

In this paper, we have derived a new MH test for minibatch MCMC methods. We demonstrated how a
simple deconvolution process allows us to use a minibatch approximation to the full data tests. We
experimentally show the benefits of our test on Gaussian mixtures and logistic regression.
Straightforward directions for future work include running more experiments with a
particular focus on investigation of the variance precondition.  More elaborate extensions include
combining our results with Hamiltonian Monte Carlo methods, providing a recipe for how to use our
algorithm (following the framework of~\cite{sgmcmc_2015}), or integrating parallel
MCMC~\cite{conf/uai/AngelinoKWSA14,conf/icml/AhnSW14} concepts.

\small
\bibliography{nips_2016}
\bibliographystyle{ieeetr}
\normalsize

\clearpage
\appendix

\begin{center}
{\Large Outline of Appendix}
\end{center}

In this appendix, we describe the following topics:

\begin{itemize}[noitemsep]
    \item The proof of Lemma~\ref{lem:worst_case}.
    \item More information on Section~\ref{ssec:gaussians}.
    \item More information on Section~\ref{ssec:logistic}.
    %\item More information on Section~\ref{ssec:nets}. % Daniel: this is in
    %progress.
\end{itemize}

\section{Proof of Lemma~\ref{lem:worst_case}}\label{app:worst_case_proof}

Assume $(\theta' - \theta) \in \pm\frac{1}{\sqrt{N}}[0.5,1]$ and $\theta -0.5
\in \pm\frac{1}{\sqrt{N}}[0.5,1]$ with matching sign. These events occur with
probability ({\color{blue} Daniel: sorry for interjecting here but don't we need
to take into account the $1/\sqrt{N}$ in the probabilities? If $N \to \infty$
(and it's going to be large with big datasets) the probabilities should be about
zero.}) $p_0=2*(\Phi(1)-\Phi(0.5)) = 0.2997$ and
$p_1=(\Phi(1)-\Phi(0.5))=0.14988$ respectively, and guarantee that
$\Lambda^*(\theta',\theta)$ is negative. This then guarantees that we can find a
$u \in [0,1]$ so that $\psi(u,\theta',\theta)$ equals
$\mathbb{E}[\Lambda^*(\theta',\theta)]$.  Specifically, choose $u_0$ to satisfy
\begin{equation}\label{eq:log_uo}
    \log u_0 = N(\Lambda^*(\theta',\theta)-\psi(1,\theta',\theta))
\end{equation}
which evaluates ({\color{blue} Daniel: sorry again for interjecting but I really
think $\log u_0$ does not have that $N$ there. Also, for the subsequent
equation, I think we need to be using \emph{expected} values of $\Lambda^*$}) to
\begin{equation}
    \log u_0 = -N(\theta'-\theta)\left(\theta-0.5+\frac{\theta'-\theta}{2}\right).
\end{equation}
This means we can rewrite the acceptance test as
\begin{equation}
    \frac{1}{b}\sum_{i=1}^b x_i-\left(\theta+\frac{\theta'-\theta}{2} + \frac{\log u_0}{N(\theta'-\theta)}\right)
    \hspace{0.1in} \not\approx \hspace{0.1in} 0
\end{equation}
where $\not\approx$ means ``significantly different from'' under the
distribution over samples of $x_i$.  The above choice of $u_0$ ensures that the
terms in parenthesis above sum to 0.5. Since the $x_i$ have mean 0.5, the test
will never correctly succeed.  Furthermore, if we sample values of $u$ near
enough to $u_0$, the terms in parenthesis will not be sufficiently different
from 0.5 to allow the test to succeed.

{\color{blue} Daniel: I would like to ask you about these probabilities because
I do not know how we got these intervals.} The choices above for $\theta$ and
$\theta'$ guarantee that
\begin{equation}
    \log u_0 \in -[0.5,1][0.75,1.5] = -[0.375,1.5]
\end{equation}
and consider the range of $u$ values
\begin{equation}
    \log u \in \log u_0 + [-0.5,0.375]
\end{equation}
the size of the range in $u$ is at least $\exp([-2,-1.125]) = [0.13534,0.32465]$ and occurs
with probability at least $p_2=0.18932$. With $u$ in this range, we rewrite the test as:
\begin{equation}
    \frac{1}{b}\sum_{i=1}^b x_i-0.5 \hspace{0.1in} \not\approx \hspace{0.1in} \frac{\log u/u_0}{N(\theta'-\theta)}
\end{equation}
so that the LHS has expected value zero.  Given our choice of intervals for the
variables, the RHS is in the range $1/\sqrt{N}[-1,0.75]$.  The standard
deviation of the LHS given the interval constraints is at least $0.5/\sqrt{b}$.
And so the gap between LHS and RHS is at most $2\sqrt{b/N}$ standard deviations.

The samples $\theta$, $(\theta'-\theta)$ and $u$ are drawn independently and so
the probability of the conjunction of these events is $c = p_0 p_1 p_2 =
0.0085$.

\section{Gaussian Mixture Experiment Details}\label{app:gaussian}

In this section, we go over the math details on the Gaussian mixture model problem borrowed
from~\cite{langevin_2011}. Our parameter is a 2-D vector $\theta =
(\theta_1, \theta_2)$, where
\begin{equation}
\theta_1 \sim \mathcal{N}(0, \sigma_1^2) \quad \mbox{ and } \quad \theta_2 \sim \mathcal{N}(0, \sigma_2^2)
\end{equation}
where $\mathcal{N}$ indicates the normal distribution (more generally, the multivariate normal). We
consider the above as our prior. Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and
$\sigma_2^2=1$, so the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
log prior probability we endow on $\theta$ is
\begin{equation}
\log p(\theta) = \log \left(\frac{1}{2\pi\sqrt{10}}\right) - \frac{1}{2}\theta^T\Sigma^{-1}\theta.
\end{equation}
To generate the data, we use the following Gaussian mixture with tied means:
\begin{equation}\label{eq:x_points}
x_i \sim \frac{1}{2}\mathcal{N}(\theta_1, \sigma_x^2) + \frac{1}{2}\mathcal{N}(\theta_1+\theta_2, \sigma_x^2)
\end{equation}
where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$. This means the log likelihood
of a single data instance is
\begin{equation}
\log p(x_i \mid \theta) = \log\left(\frac{1}{4\sqrt{\pi}}\right) +
\log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
(\theta_1+\theta_2))^2\right)\right)
\end{equation}
Here is the problem statement: given some number of conditionally independent data points $x_1, x_2,
\ldots, x_N$ generated according to~(\ref{eq:x_points}), determine the posterior
distribution of $\theta$:
\begin{equation}\label{eq:log_post}
\log p(\theta \mid x_1,\ldots,x_N) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
\end{equation}
Alternatively, if there are too many data points, we may opt to instead pick a point estimate of
$\theta$, generally the MAP estimate. (If $N$ is extremely large, it will cause the posterior to
peak sharply at its modes, reducing distribution estimates to point estimates.) Note that in many
cases, we will need to take a \emph{minibatch estimate} of~(\ref{eq:log_post}). In that case, the
literature generally uses
\begin{equation}\label{eq:scaling_factor}
\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
\end{equation}
where we only use $n \ll N$ samples, but we must scale up the likelihood contribution by $N/n$. If we
didn't add this scaling factor, then the contribution of the likelihood terms would be weaker.

One technique we use is adding a \emph{temperature} to our distribution. In general, we will want to
add $T > 1$ so that our posterior is $p(\theta)((\prod_{i=1}^n p(x_i\mid \theta))^{N/n})^{1/T}$,
resulting in the log posterior of 
\begin{equation}\label{eq:log_prior_temp}
\log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
\end{equation}
which has the extra $1/T$ to decrease the scale factor. Equation~(\ref{eq:log_prior_temp}) is what
we use for our experiments, because warmer distributions help us satisfy our ${\rm std}(\Delta') <
1.2$ requirement.

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{contour_v1}
  \caption{The posterior distribution, from 50 to 100k samples, with temperature set at 1.}
  \label{fig:contour1}
\end{figure}
\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{contour_v2}
  \caption{The posterior distribution, with $N=10000$ but with temperature $T$ varying.}
  \label{fig:contour2}
\end{figure}

To gain some intuition on what the posterior looks like, Figure~\ref{fig:contour1} shows simulated
contour plots of the posterior based on varying numbers of data points $N$, with the temperature set
at $T=1$. Note that because we are using all $N$ points here, the scale factor $N/n=1$. As $N$
increases, the posterior converges to a multimodal distribution with modes at $(0,1)$ and $(1,-1)$.
Figure~\ref{fig:contour2} is similar, except this time we fix the number of samples at $N=10000$,
but show how changing the temperature $T$ affects the distribution. A larger $T$ implies a flatter
posterior, one that (weakly) peaks in between the two true modes.

\section{Logistic Regression Experiment Details}\label{app:logistic}

In this section, we go over some details of our logistic regression experiment. The feature vector
for an image consists of its pixel values, normalized between 0 and 1. For simplicity, we only
consider the binary classification case, so we only use digits one (denoted as output $y = 1$) and
seven (denoted as $y = -1$). The probability of the $i^{\rm th}$ output $y_i \in \{-1,+1\}$ with the
input vector $x_i$ is
\begin{equation}
p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)},
\end{equation}
where $\theta$ is the 784-length parameter vector.

For our experiment, we impose a uniform prior to represent our lack of knowledge about $\theta$.  We
use a random walk proposer, which can be modeled as $\theta' = \theta_i + \mathcal{N}(0,
\sigma^2I)$, where $\theta_i$ is the current sample, $\theta'$ is the proposed sample, and we choose
the variance to be a constant $\sigma^2 = 0.05$ for all components. We initialize $\theta_0$ to be a
vector of all ones, and set our minibatch size as $m=100$.

For our minibatch MH test, in order to enforce the ${\rm std}(\Delta') < 1.2$ condition, we use a
constant temperature $T = 1000$. If our estimated ${\rm std}(\Delta') \ge 1.2$, we ignore the
current iteration.  %Figure~\ref{fig:appendixExp2} plots our estimated ${\rm std}(\Delta')$ values
%versus iteration count.

%\begin{figure}[t]
  %\centering
  %\includegraphics[width=1\linewidth]{./figures/appendixExp2}
 % \caption{Additional results for the logistic regression experiment.}
  %\label{fig:appendixExp2}
%\end{figure}

%For adaptive MH testing, our experimental settings are the same as with our MH test, except we do
%not impose a temperature. The minibatch size of adaptive MH testing is also initialized as 50, but
%it may increase by that amount each iteration. Figure~\ref{fig:appendixExp2} shows the histogram of
%the actual minibatch size at the end of each iteration in adaptive MH testing.

\iffalse
\section{Neural Network Experiment Details}\label{app:nnet}

For our neural network experiment, we use the architecture discussed in Section~\ref{ssec:nets}. We
use the SGHMC as the proposer for our minibatch MH testing. For the baseline, we use a tuned
adaptive gradient descent optimizer, whose step size changes by $0.01 /(i+1)^{0.4}$. Both methods
have minibatch sizes set at $m=200$. There are one million total data instances $x_i$.

For SGHMC, we use the simplified update equations~\cite{sghmc_2014}:
\begin{equation}
\Delta \theta = v, \quad \quad\Delta v = -\eta \nabla U(\theta) - \alpha v + \mathcal{N}(0, 2(\alpha
-\hat{\beta}) \eta),
\end{equation}
where $v$ represents auxiliary momentum variables, and $\nabla U(\theta)$ is the gradient of the
system. We set the hyperparameters to be $\eta_i = 0.01 /(i+1)^{0.4}$, and $\alpha=0.1$. We use the
empirical Fisher~\cite{conf/icml/AhnBW12} information $V(\theta)$ to estimate the value of
$\hat{\beta}$, so that $\hat{\beta}=\frac{1}{2}\eta V(\theta)$. In order to control ${\rm
std}(\Delta')$, we initialize the temperature at 1000, and adjust it at iteration $i$ according to
$T_i = \max\{1, 1000/(i+1)^{0.5}\}$.

\fi

% Daniel: here are example LaTeX codes for figures and tables if we want to use them.
%\begin{figure}[h]
%  \centering
%  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
%  \caption{Sample figure caption.}
%\end{figure}
%\begin{table}[t]
%  \caption{Sample table title}
%  \label{sample-table}
%  \centering
%  \begin{tabular}{lll}
%    \toprule
%    \multicolumn{2}{c}{Part}                   \\
%    \cmidrule{1-2}
%    Name     & Description     & Size ($\mu$m) \\
%    \midrule
%    Dendrite & Input terminal  & $\sim$100     \\
%    Axon     & Output terminal & $\sim$10      \\
%    Soma     & Cell body       & up to $10^6$  \\
%    \bottomrule
%  \end{tabular}
%\end{table}
%\usepackage[pdftex]{graphicx} ...
%\includegraphics[width=0.8\linewidth]{myfile.pdf}

\end{document}