### 2017-03-29 old_paper_versions/NIPS_Reviews_Rebuttals/rebuttal_reviewers.txt

We thank the reviewers for their thoughtful commentary on our paper.
We can only agree with the comments and appreciate the reviewers
patience in reviewing the paper. It was an early draft. We noticed
most of these issues ourselves in the time since we submitted the paper
and have been working to address them. We discuss those corrections
in detail below.

General:

First we established worst-case lower bounds for references [3] and [4]
of Omega(N) bounds on the number of samples consumed per iteration,
using simple Gaussian examples.

In addition, Section 4 (the theoretical results) has been
substantially revised.  Rather than general appeal to CLT, we include
bounds based on t-statistics (using both the empirical mean and
variance) asymptotics from [A] (below) with explicit constants and an
O(n^-0.5) error bound. We revised the analysis of the test error rate
which now requires no additional assumptions. We include error bounds
on the equilibrium distribution given bounds on transition kernel
error which appeared in [3] and [4], citing those works, under the
appropriate assumptions.

To Reviewers 2, 3, and 4 regarding the Barker function:

You are correct, the logistic (Barker) function is less efficient than
the Metropolis acceptance test with \theta' close to \theta when run
on the full dataset. However, there is no comparable result when these
tests are used on minibatches since they are applied in completely
different ways. Empirically, we find that we can scale proposal step
size or temperature to achieve as small minibatch size as desired
using our corrected Barker test, but this is not true of [3] or [4]
which via the analysis above (and in experiments) still require
Omega(n) samples for some single tests. Thus the approx.  Barker test
is arbitrarily more efficient than tail tests in terms of average data
consumed per generated distinct sample. We have revised the
experiments so show that such gaps are up to 3 orders of magnitude for
logistic regression and the gaussian mixture model.

To Reviewers 1, 2, and 3 regarding the deconvolution (in)stability:

We now describe in detail the construction of the distribution of
X_corr. We explored linear programming and least squares solutions. We
find that the latter allow us to compute finer discrete approximations
to the distribution of X_corr (4000 points). It also gives smaller
absolute (i.e. L1) error. Even though least squares minimizes the L2
norm of the error, the L1 error is also bounded down to 1e-5 by L2
norm minimization. The solution accuracy benefits from fairly strong
L2 regularlization, which also makes the inversion
(i.e. deconvolution) more stable. To ensure positivity of all three
distributions, we impose a stricter bound on the variance of X_norm.
This effectively shifts variance to $X_corr$ by convolving it with
another normal distribution, and guarantees that it is positive.
We computed X_corr with variance bounds of 0.8 and 0.9 both of
which give CDF errors below 1e-4.

To Reviewer 3:

(TODO respond to experiment criticism, but I really can't defend the work as it is ...)

Both Reviewers 1 and 3 correctly identified our implicit assumptions
about Lemma 2. Reviewer_1 is right that we approximated it for large
N, so sampling without replacement approximates sampling with
replacement.  In the case of very large datasets however, it is also
typically more reasonable to treat the dataset as itself a sample
drawn from a still larger, or possibly infinite, dataset (e.g. this
holds for all the examples in our paper), and the instance
"resolution" implies that repeats have probability ~ zero. So
distinctions between sampling with replacement and without are perhaps
moot for these datasets.

To Reviewers 1 and 2:

Thank you for mentioning [BDH], "MCMC Methods for Tall Data." We added
this reference as part of our ongoing revision of the paper and will
substantially address issues raised in [BDH]. In addition, we are
addressing Reviewer 2's concern regarding the difference between ours
and [BDH] Section 6.3. The idea of using a Barker test with minbatches
was proposed in [BDH]. However, a crucial addition in our approach is
the use of an additive correction variable. Thus in our approach,
there is no lower bound on error in our approximate test, and we show
it decreases with O(n^-0.5). In [BDH] there is no correction between
normal and logistic CDFs, and the method has a minimum error set by
the max absolute difference between their CDFs. Another difference is
that the idea proposed in [BDH] was not explored experimentally. In
particular it was not shown that tests can be done with arbitrarily
small minibatches in [BDH].

To Reviewer 7:

We have improved the algorithm to use an adaptive step size rather
than fail on some tests.  Unlike [3] and [4] however, since the amount
of data required for our test scales in proportion to the proposal
step size, the minibatch size has a normal size bound.

[A] Y. Novak, "On Self-Normalized Sums And Student’s Statistic"
Theory of Probability and its Applications, 49(2), 336-344, 2005.